# Intermediate R

<https://learn.datacamp.com/courses/intermediate-r>

<span style="color:teal;font-weight:bold">(Note: If you do [Intermediate R for Finance](https://learn.datacamp.com/courses/intermediate-r-for-finance) instead, change the title of this bookdown chapter above and change the DataCamp chapter titles below. Whichever version of Intermediate R you do, delete this note)</span>




## Conditionals and Control Flow
Relational Operators :--

Operator  |  Used for  |
:---------:  |  :---------:  |
<  |  Less Than | 
> | Greater Than  |  
<=  | less than or equal to  |  
>=  |  greater than or equal to |  
==  | Equal To |
!=  | Not Equal To |

Logical Operators :-- 

Operator  |  Used for  |   
:---------:  |  :---------:  |
&  | (AND) passess TRUE If both the conditons are TRUE | 
|  |   (OR) passes TRUE if anyone of the consitions are TRUE |
!     | (NOT) Changes the logical value from TRUE to FALSE and otherwise  |  

We can use these relational operators for vectors, matrices and data frames. 

```{r}
numbers1 <- c(78,90,4,35,65)
numbers2 <- c(57,65,876,234,45)
numbers1>50
numbers2>=numbers1
```
Use of logical operators :--
```{r}
numbers1 >=50 | numbers1<=100
lows <- numbers1 >=10 & numbers1 <= 50
lows


#sum of the extremes numbers 
sum(lows)

#Use of not !
!lows 

```
Conditional Statement :-- 

The structure of conditional statement is :--

<span style="color:red"> if(condition) </span> {  
execute if condition TRUE  
}<span style="color:red"> else if(condition) </span> {   
execute when if condition is FALSE and else if condition TRUE  
}<span style="color:red"> else </span> {  
execute when if and else if condition is FALSE  
}

We can also use conditional statements inside a conditional statement. 

```{r}
Income <- 70000
if (Income < 40000){
  if(Income<20000){
    print("Need Financial Aid")
  }else{
    print("Fine enough")
  }
} else if (Income< 80000){
  print("Good standards of living")
} else {
  print("rich")
}

```


## Loops
Loops can come in handy on numerous occasions. While loops are like repeated if statements, the for loop is designed to iterate over all elements in a sequence.  The simple syntax for while loop is :-- 
<span style="color:red"> while(condition) </span> {  
execute if condition TRUE  and keep iterating until condition is FALSE
}
```{r}
course_student <- 0 
while(course_student <= 23){
  print(paste("Total Stdent in class is now ", course_student))
  course_student <- course_student+1
}
```
The Break statement:--

The **break** statement breaks out of the while loop whenever a certain condition is made.

The other statement to continue the loop further is **next**

For Loop:

We can use a for loop to print all the elements in a vectors, matrices, lists, data frames.There are two ways create a for loop. Both of them creates same output. But using the second type we can control the indexing of the vector or list.

```{r}
courses <- list("Discrete Math", "Calculas", "Complex Sequences", "PRobability","Combinatorics")
#First way to create for loop
for(c in courses){
  print(c)
}
#Second way of for loop
for(i in 1:length(courses)){
  print(courses[[i]])
}
```
Nested for loop :--
```{r}
#We are using the matrix we created before in the first chapter

Firstyear_profit <- c(340,890)
Secondyear_profit <- c(900,234)
Thirdyear_profit <- c(300,450)

Business_matrix_new <- matrix(c(Firstyear_profit, Secondyear_profit, Thirdyear_profit), byrow=TRUE, nrow = 3)
for (i in 1:nrow(Business_matrix_new)) {
  for (j in 1:ncol(Business_matrix_new)) {
    print(paste("On row", i, "and column", j, "the matrix contains", Business_matrix_new[i,j]))
  }
}
```
We can also use if and else statement inside the for loop. 
```{r}
number <- c(2,3,4,7,9,11,13,17,19,23,29,31)
for (n in number) {
  if (n > 15) {
    print(paste(n, "is greater than 15"))
  } else {
    print(paste(n, "is less than 15"))
  }
  print(n)
}
```

## Functions
Functions are really important in R do a lot of works . There are a lot of functions in a lot of packages on R. Later in this chapter we will figure out how we can load the packages in R. 
To consult for the documentation of any function we can use the below commands which will prompt to the R website and show us the documentation and we can also look for the required arguments needed for the functions.

```{r}
#Seeing the documentation for the sd function 
help(sd)
?sd
#To see the arguments of the sd function
args(sd)
```
A use of the mean function is shown below :--
```{r}
value1 <- c(16, 9, 13, 5, NA, 17, 14)
value2 <- c(17, NA, 5, 16, 8, 13, 14)
avg_sum <- mean((value1+value2), na.rm = TRUE)
avg_sum_trimmed <- mean((value1+value2), trim = 0.2, na.rm = TRUE)
avg_sum_trimmed
avg_sum
```
We can also create our own function using the format :--  
first_function <- <span style="color:red"> function(arg1,arg2) </span> {  
body
}   
```{r}
cube_function <- function(arg){
  cubic <- arg^3
  return(cubic)
}
cube_function(3)
```
We can also make functions with no arguments.

```{r}
function_witoutarg <- function(){
  print("This is a function without any argument")
}
function_witoutarg()
```
Function Scoping:

Function scoping implies that variables that are defined inside a function are not accessible outside that function.
We can also add control flow statement in the function body

```{r}
linkedin <- c(16, 9, 13, 5, 55, 17, 14)
facebook <- c(17, 2, 5, 16, 8, 13, 14)

interpret <- function(num_views) {
  if (num_views > 15) {
    print("You're popular!")
    return(num_views)

  } else {
    print("Try to be more visible!")
    return(0)

  }
}
interpret(linkedin[1])

```
No we should know how can we load and install R packages:--

We can use **install.packages("")** to install the packages first and then to load packages we can use **library()** Then we can see the packages we have loaded using **search()**

```{r}
install.packages("tidyverse")
library(tidyverse)

```
```{r}
search()
```



## The apply family
The syntax of lapply is lapply(Vector, function). It takes a vector list and applies a function to each of its members.

```{r}
digits <- c(16, 9, 13, 5, 55, 17, 14)
cubics <- lapply(digits, cube_function)
cubics
```
We can also use annonymous function with lapply. 
```{r}
square <- lapply(digits, function(x){x*x})
square
```

sapply() function does the same job as lapply() function but returns a vector. Like lapply, we can use our own functions for sapply.

```{r}
sapply(digits, cube_function)
```
There is another way of using dunction in data that is vapply. The syntax is :--

<span style="color:red"> vapply(X, FUN, FUN.VALUE, …, USE.NAMES = TRUE) </span>
Over the elements inside X, the function FUN is applied. The FUN.VALUE argument expects a template for the return argument of this function FUN. USE.NAMES is TRUE by default; in this case vapply() tries to generate a named array, if possible.

```{r}
normal <- function(x){
  c(min = min(x),max= max(x), mean= mean(x))
}
vapply(digits, normal, numeric(3))
```
Summary:

*  lapply():  
   +  apply function over list or vector  
   +  output = list. 
*  sapply():  
   +  apply function over list or vector 
   +  try to simplify list to array  
*  vapply():  
     +  apply function over list or vector  
     +  explicitly specify output format.  
     
## Utilities
Mathematical Utilities:  

*  abs(): Calculate the absolute value.   
*  sum(): Calculate the sum of all the values in a data structure.  
*  mean(): Calculate the arithmetic mean.  
*  round(): Round the values to 0 decimal places by default.  
```{r}
digits <- c(16, 9, 13, 5, 55, 17, 14)
sum(round(abs(digits)))
vec1 <- c(2.3,4.5,1.2,3.6,9.3)
vec2 <- rev(vec1)

mean_vec_abs <- mean(c(abs(vec1), abs(vec2)))
mean_vec_abs
```


*  seq(): Generate sequences, by specifying the from, to, and by arguments.   
*  rep(): Replicate elements of vectors and lists.   
*  sort(): Sort a vector in ascending order. Works on numerics, but also on character strings and logicals.   
*  rev(): Reverse the elements in a data structures for which reversal is defined.  
*  str(): Display the structure of any R object.   
*  append(): Merge vectors or lists.   
*  is.(): Check for the class of an R object.  
*  as.(): Convert an R object from one class to another.  
*  unlist(): Flatten (possibly embedded) lists to produce a vector.


```{r}
linkedin <- list(16, 9, 13, 5, 2, 17, 14)
facebook <- list(17, 7, 5, 16, 8, 13, 14)

# Convert linkedin and facebook to a vector: li_vec and fb_vec
li_vec <- unlist(linkedin)
fb_vec <- unlist(facebook)

# Append fb_vec to li_vec: social_vec
social_vec <- append(li_vec, fb_vec)

# Sort social_vec
sort(social_vec, decreasing = TRUE)
```

```{r}
rep_seq <- rep(seq(1, 5, by = 2), times = 7)
rep_seq
#sequance from 1 to 100
seq(1, 100, by = 2)
```
Regular Expressions:

Regular expressions are used to clean data to get them ready to work with after we get them. They can be used to see whether a pattern exists inside a character string or a vector of character strings.

[GREP,GREPL]

grepl(), which returns TRUE when a pattern is found in the corresponding character string. grep(), which returns a vector of indices of the character strings that contains the pattern.

```{r}
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")
edu_has <- c(grepl("edu", x = emails))
edu_emails <- emails[edu_has]
edu_emails
```
We can use the following to match contents in a string :  

*  we cam ise the caret, ^, and the dollar sign, $ to match the content located in the start and end of a string, respectively.  

*   @, because a valid email must contain an at-sign.  
*  (.*), which matches any character (.) zero or more times. Both the dot and the asterisk are metacharacters. We can use them to match any character between the at-sign and the “.edu” portion of an email address.    

*  \.edu$, to match the “.edu” part of the email at the end of the string. The \ part escapes the dot: it tells R that you want to use the . as an actual character.  


```{r}
# Use grepl() to match for .edu addresses more robustly
grepl("@.*\\.edu$", x = emails)
```
```{r}
# Use grep() to match for .edu addresses more robustly, save result to hits
hits <- grep("@.*\\.edu$", x = emails)

# Subset emails using hits
emails[hits]
```

[SUB,GSUB]

**sub()** and **gsub()** can be used to replace a character within a element. sub() only replaces the first match, whereas gsub() replaces all matches.
```{r}
sub("@.*\\.edu$", "@datacamp.edu", emails)
```

Date and Times:
```{r}
today_date <- Sys.Date()
now <-  Sys.time()
today_date
```

We can also do date maths like difference of dates 
```{r}
Newdate <- as.Date("2001-03-09")
dif<- today_date-Newdate
dif
```

We cam use the following formats to change, create dates using as.Date(). Date increments in Days.

*  %Y: 4-digit year (1982)   
*  %y: 2-digit year (82)   
*  %m: 2-digit month (01)   
*  %d: 2-digit day of the month (13)   
*  %A: weekday (Wednesday)   
*  %a: abbreviated weekday (Wed)   
*  %B: month (January)   
*  %b: abbreviated month (Jan)  

```{r}
 #Definition of character strings representing dates
str1 <- "May 23, '96"
str2 <- "2012-03-15"
str3 <- "30/January/2006"

# Convert the strings to dates: date1, date2, date3
date1 <- as.Date(str1, "%b %d, '%y")
date2 <- as.Date(str2, "%Y-%m-%d")
date3 <- as.Date(str3, "%d/%B/%Y")


# Convert dates to formatted strings
format(date1, "%A")
format(date2, "%d")
format(date3, "%b %Y")
```
We cam use the following formats to change, create time using as.POSIXct(). Time increments in Seconds.  


*  %H: hours as a decimal number (00-23)   
*  %I: hours as a decimal number (01-12)   
*  %M: minutes as a decimal number   
*  %S: seconds as a decimal number  
*  %T: shorthand notation for the typical format %H:%M:%S   
*  %p: AM/PM indicator

```{r}
# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2)

# Convert times to formatted strings
format(time1, "%M")
format(time2, "%I:%M %p")
```








