# Intro to Data Visualization with ggplot2

<https://learn.datacamp.com/courses/introduction-to-data-visualization-with-ggplot2>

```{r}
## Load all packages used in this chapter
library(tidyverse)
library(ggplot2)
```


## Introduction
Previously in the chapters we have seen how to create a plot using ggplot(). In this chapter we will explore the functions we can use to make our plot more understandable and easy to read. 
 Data visualization is really important skill in the data science field and we need to design our plots nicely to achieve the best from the statistics. 
 
We can change a numerical or categorical value as factor by using the function  <span style="color:red">factor()</span> :--

```{r}
summary(mtcars)

ggplot(mtcars, aes(factor(cyl), mpg)) + geom_point()

```
  

Jargon of each element :--  

Data  |  Used for  |  
:---------:  |  :---------:  |  
Aesthetics  |  x-axis,y-axis,color,fill,size,label,alpha,shape,line width,linetype |  
Geometrics | point, line, histogram, bar, boxplot  |   
Themes  | non-data ink  |   
Statistics  |  binning, smoothing, descriptive, inferential |   
Coordinates  | cartesian, fixed, polar, limits | 
Facets | columns, rows |  


The grammatical elements description :--   

*  Data (Data set being plotted)  
*  Aesthetics (The scales onto which we map our data)  
*  Geometries (The visual elements used for our data)  
*  Themes(All non-data ink)     
*  Statistics (Representations of our data to aid understanding)  
*  Coordinates ( The space on which data will be plotted)  
*  Facets (Plotting small multiples)    


We can use the <span style="color:red">geom_smooth()</span> layer to look for patterns in our data. We use this layer to Plot two continuous position variables in the graph.
We can also use an argument called **alpha=** in the geom_point() to set this to a point which is the opacity of the point. If we input 0 as alpha it will show all the points as transparent.

```{r}
str(diamonds)
ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  geom_smooth() 
```

```{r}
ggplot(diamonds, aes(carat, price, color = clarity)) +
  geom_point(alpha = 0.4) +
  geom_smooth()
```

## Aesthetics


These are the aesthetics you can consider within aes() :--  x, y, color, fill, size, alpha, labels and shape.(in order). 
```{r}
head(iris)
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width,  size= Petal.Width, color= Species)) +
  geom_point(alpha = 0.5) +
  geom_smooth()
```


**Remember that shape and label aesthetics are only for categorical data**  

Typical visible aesthetics:--  

*  x = X axis position.  
*  y = Y axis position.  
*  fill = Fill color.   
*  size = Are or radius of points, thickness of lines.   
*  alpha = Transparency.   
*  linetype = line dash pattern.   
*  labels = Text on a plot or axes.   
*  shape = Shape.   

Attributes:--  
We use aesthetics in the aes() and we use attribute in the geom_*() . 

```{r}
# A hexadecimal color
my_blue <- "#4ABEFF"

# Change the color mapping to a fill mapping (AESTHETICS)
ggplot(iris, aes(Sepal.Length, Sepal.Width, color= Species)) +
  
  # Set point size and shape (ATTRIBUTES)
  geom_point( size = 4, shape = 1)
```

In order to add text, we use the geom_text() function and to use row names of a data set, we set the label of the attribute - label as rownames(name of data set).  

```{r}
ggplot(mtcars, aes(wt, mpg, color = fcyl)) +
  # Add text layer with label rownames(mtcars) and color red
  geom_text(label = rownames(mtcars), color = 'red')

```

Modifying Aesthetics:--  

Position: 
Adjustment for overlapping :--   

*  identity.  
*  dodge  
*  stack.  
*  fill  
*  jitter  
*  jitterdodge  
*  nudge  

The default for position is the **"identity"**. The **"jitter"** maintains consistency across plots and layers. 
```{r}
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, color= Species)) + geom_point(position = "jitter")
```

Scale functions :--  

*  scale_x_*()  
*  scale_y_*()  
*  scale_color_*()  
*  scale_fill_*()  
*  scale_shape_*()  
*  scale_linetype_*()  
*  scale_size_*()  

We can set the labels for x and y axis which is really important to know which axis is used to plot which variable and describe it. we use <span style="color:red">labs()</span> to set the x- and y-axis labels. It takes strings for each argument.  
<span style="color:red">scale_color_manual()</span> defines properties of the color scale (i.e. axis). The first argument sets the legend title. values is a named vector of colors to use.  

To Implement a custom fill color scale we use scale_fill_manual(). 

```{r}
palette <- c(automatic = "#377EB8", manual = "#E41A1C")

# Set the position
ggplot(mtcars, aes(x=factor(cyl), fill = factor(am))) +
  geom_bar(position="dodge") +
  labs(x = "Number of Cylinders", y = "Count")
  scale_fill_manual("Transmission", values = palette)
```
You can make one variable plots in ggplot2, but you will need to add a fake y axis by mapping y to zero. 
When using setting y-axis limits, you can specify the limits as separate arguments, or as a single numeric vector. That is, ylim(lo, hi) or ylim(c(lo, hi)). 
```{r}
ggplot(mtcars, aes(mpg, 0)) +
  geom_jitter() +
 ylim(-2,2)
```


## Geometries
Scatter plots (using geom_point()) are intuitive, easily understood, and very common, but we must always consider overplotting, particularly in the following four situations:  

*  Large datasets  
*  Aligned values on a single axis  
*  Low-precision data  
*  Integer data  

Typically, alpha blending (i.e. adding transparency) is recommended when using solid shapes. Alternatively, you can use opaque, hollow shapes.

Small points are suitable for large datasets with regions of high density (lots of overlapping). 
Overplotting 2: Aligned values

This occurs when one axis is continuous and the other is categorical, which can be overcome with some form of jittering.
```{r}
plt_mpg_vs_cyl_by_am <- ggplot(mtcars, aes(cyl, mpg, color = am))

# Default points are shown for comparison
plt_mpg_vs_cyl_by_am + geom_point()
```
```{r}
# Now jitter and dodge the point positions
plt_mpg_vs_cyl_by_am + geom_point(position = position_jitterdodge(jitter.width = 0.3, dodge.width = 0.3))
```
Let's take a look at another case:

Low-precision data
This results from low-resolution measurements like in the iris dataset, which is measured to 1mm precision (see viewer).
```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  # Use a jitter position function with width 0.1
  geom_point(alpha = 0.5, position=position_jitter(width=0.1))
```

Histograms:--  

*  Always set meaningful binwidth for histograms.   
*  No spaces between bars   
*  X axis labels are between bars   


Positions of Histograms:--  

*  stack (the default): Bars for different groups are stacked on top of each other.    
*  dodge: Bars for different groups are placed side by side.     
*  fill: Bars for different groups are shown as proportions.    
*  identity: Plot the values as they appear in the dataset.   

```{r}
ggplot(iris, aes(x= Sepal.Width, fill= Species)) + geom_histogram(binwidth = 0.1, center=0.05)
```

```{r}
ggplot(iris, aes(x= Sepal.Width, fill= Species)) + geom_histogram(binwidth = 0.1, center=0.05, position = "stack")
```

```{r}
ggplot(iris, aes(x= Sepal.Width, fill= Species)) + geom_histogram(binwidth = 0.1, center=0.05, position = "dodge")
```

```{r}
ggplot(iris, aes(x= Sepal.Width, fill= Species)) + geom_histogram(binwidth = 0.1, center=0.05, position = "fill")
```

Bar Plots:

*  geom_bar() [stat = “count”] : counts the number of cases at each X position    
*  geom_col() [stat = “identity”] : plots actual values   

```{r}
ggplot(mtcars, aes(cyl, fill = am)) +
  # Add a bar layer
  geom_bar()
```

```{r}
ggplot(mtcars, aes(cyl, fill = am)) +
  # Change the position to "dodge"
  geom_bar(position = "dodge")
```

```{r}
ggplot(mtcars, aes(cyl, fill = am)) +
  # Set the position to "fill"
  geom_bar(position = 'fill')
```


Line Plots:--  


```{r}
str(economics)
ggplot(economics, aes(date, unemploy/pop)) +geom_line()

```


## Themes
Themes are all non-data ink visual elements which are not part of the data. THere are three types of themes :--   

*  Text : element_text()     
*  Line : element_line()  
*  Rectangle : element_rect()  

Legend :-- A legend of a plot helps to understand which series or groups corresponds to each bar, line, box or observations, based on its type, color or both.  


To change stylistic elements of a plot, call theme() and set plot properties to a new value. For example, the following changes the legend position.  

[p + theme(legend.position = new_value)]. 

Here, the new value can be. 

*  “top”, “bottom”, “left”, or “right’”: place it at that side of the plot.  
*  “none”: don’t draw it.  
*  c(x, y): c(0, 0) means the bottom-left and c(1, 1) means the top-right.  

```{r}
ggplot(diamonds, aes(carat, price, color = clarity)) +
  geom_point(alpha = 0.4) +
  geom_smooth() +
  theme(
    # For all rectangles, set the fill color to grey92
    rect = element_rect(fill = "grey92"),
    # For the legend key, turn off the outline
    legend.key = element_rect(color = NA)
  )
```
Modifying whitespace:--  

Whitespace means all the non-visible margins and spacing in the plot.

To set a single whitespace value, use unit(x, unit), where x is the amount and unit is the unit of measure.

Borders require you to set 4 positions, so use margin(top, right, bottom, left, unit). To remember the margin order, think TRouBLe.

The default unit is "pt" (points), which scales well with text. Other options include "cm", "in" (inches) and "lines" (of text).   


```{r}
ggplot(mtcars, aes(wt, mpg , color= factor(cyl))) + geom_point() +theme(
    # Set the axis tick length to 2 lines
    axis.ticks.length = unit(2, "lines")
  )
```
```{r}
ggplot(mtcars, aes(wt, mpg , color= factor(cyl))) + geom_point() + theme(
    # Set the legend margin to (20, 30, 40, 50) points
    legend.margin = margin(20,30, 40, 50, "pt")
  )
```
```{r}
ggplot(mtcars, aes(wt, mpg , color= factor(cyl))) + geom_point() + theme(
    # Set the plot margin to (10, 30, 50, 70) millimeters
   plot.margin = margin(10,30,50,70, "mm")
  )
```


Defining Theme objects :--   

*  Useful when you have many plots  
*  Provides consistency in style  
*  Apply a specific theme everywhere  

```{r}

new <-
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species))+
geom_jitter (alpha = 0.6) +
scale_x_continuous("Sepal Length (cm)", limits = c(4,8), expand = c(0,0)) +
scale_y_continuous ("Sepal Width (cm)", limits = c(1.5,5), expand = c(0, 0))+
scale_color_brewer("Species", palette = "Dark2", labels = c ("Setosa",
"Versicolor","Virginica"))

new + theme(text = element_text (family="serif", size = 14),
            rect = element_blank (),
            panel.grid = element_blank() ,
            title = element_text(color="#8b0000"),
            axis.line = element_line(color ="black"))
```
We can use this theme by saving it to a variable and using it in any other plots. 
```{r}
theme_new <- theme(text = element_text (family="serif", size = 14),
            rect = element_blank (),
            panel.grid = element_blank() ,
            title = element_text(color="#8b0000"),
            axis.line = element_line(color ="black"))

diamond_plot <- ggplot(diamonds, aes(carat, price, color = clarity)) +
  geom_point(alpha = 0.4) +
  geom_smooth() 
diamond_plot + theme_new

```
We can aslo access to some built-in themes like theme_classic() and theme_bw() 
```{r}
new + theme_bw()
```
```{r}
#Access library gg themes for more themes
library(ggthemes)
diamond_plot + theme_tufte()
```

Effective Explanatory Plots:--  

```{r}
library(gapminder)
gm2007_complete <- gapminder %>% 
  filter(year == 2007) %>% 
  select(country, lifeExp, continent) %>% 
  arrange(lifeExp)
gm_2007_asia <- gm2007_complete %>% filter(continent =="Asia")
gm_2007_asia
```
```{r}
ggplot(gm_2007_asia, aes(x = lifeExp, y = country, color = lifeExp)) +
  geom_point(size = 4) +
  geom_segment(aes(xend = 30, yend = country), size = 2) +
  theme(legend.position="right")
```
```{r}
ggplot(gm_2007_asia, aes(x = lifeExp, y = country, color = lifeExp)) +
  geom_point(size = 4) +
  geom_segment(aes(xend = 30, yend = country), size = 2) +
  theme(legend.position="right")+
  geom_text(aes(label = lifeExp), color = "white", size = 1.5)
```

```{r}


# Add a title and caption
ggplot(gm_2007_asia, aes(x = lifeExp, y = country, color = lifeExp)) +
  geom_point(size = 4) +
  geom_segment(aes(xend = 30, yend = country), size = 2) +
  geom_text(aes(label = round(lifeExp,1)), color = "white", size = 1.5) +
  scale_x_continuous("", expand = c(0,0), limits = c(30,90), position = "top") +
 
 labs(title="Highest and lowest life expectancies, 2007", caption="Source: gapminder")
```
We can use nannotate() to add text and a curve to the plot.

```{r}
diamond_plot +geom_vline(xintercept = 3, color = "grey40", linetype = 3) +
  annotate(
    "text",
    x = 3.5, y = 4900,
    label = "The\nrandom\nline",
    vjust = 1, size = 3, color = "grey40"
  ) +
  annotate(
    "curve",
    x = 3.5, y = 4900,
    xend = 3, yend = 5200,
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    color = "grey40"
  )
```
```{r}
global_mean <- mean(gm2007_complete$lifeExp)
x_start <- global_mean + 4
y_start <- 5.5
x_end <- global_mean
y_end <- 7.5
ggplot(gm_2007_asia, aes(x = lifeExp, y = country, color = lifeExp)) +
  geom_point(size = 4) +
  geom_segment(aes(xend = 30, yend = country), size = 2) +
  geom_text(aes(label = round(lifeExp,1)), color = "white", size = 1.5) +
  scale_x_continuous("", expand = c(0,0), limits = c(30,90), position = "top") +
 
 labs(title="Highest and lowest life expectancies, 2007", caption="Source: gapminder") +
  geom_vline(xintercept = global_mean, color = "grey40", linetype = 5) +
  annotate(
    "text",
    x = x_start, y = y_start,
    label = "The\nglobal\naverage",
    vjust = 1, size = 3, color = "red")+
  annotate(
    "curve",
    x = x_start, y = y_start,
    xend = x_end, yend = y_end,
    arrow = arrow(length = unit(0.2, "cm"), type = "closed"),
    color = "grey40")
```

