# Data Manipulation with dplyr

<https://learn.datacamp.com/courses/data-manipulation-with-dplyr>

```{r}
library(dplyr)
library(readr)
library(tidyverse)
library(tidycensus)
library(gapminder)
```


##  Transforming Data with dplyr
We can use a lot of verbs from dplyr to select rows , columns and we can also filter the data and we can also arrange the data in an ascending or a descending order. 
The functions are :--  select(),filter(),arrange(),mutate()

We can use select() function to select particular variables from the data
```{r}
head(gapminder)
T1 <- gapminder %>% select(country,year,continent,pop,lifeExp)
```
We can use the filter() function and use a condition as an argument and filter our dataset 
```{r}
filtered <- gapminder %>% filter(year==1952)
head(filtered)
```
We can use arrange() function and order the datset according to an ascending order we can also use desc() as an argument in the arrange() function and order it in a descending order
```{r}
arranged <- filtered %>% arrange(desc(gdpPercap))
head(arranged)
```
We can use mutate() to create a new variables or we can change an existing one 

```{r}
T1 %>% mutate(total_lifeExp_inMill = (lifeExp * pop)/1000000) %>% arrange(desc(total_lifeExp_inMill))
head(T1)
```




## Aggregating Data
We can use the count() function to count the number of observations. We can use count and sort together to count and then arrange. We can use the sort=TRUE argument to sort it and we can also use the wt= arguamnt to wight it according to a certain variable

```{r}
T1 %>% count(continent, sort=TRUE, wt= mean(lifeExp))
```

We can then use summarize() to summarize the data in groups by a specific variable

```{r}
T1 %>% 
  group_by(country) %>% 
  summarize(mean_lifeExp = mean(lifeExp)) %>% 
  arrange(desc(mean_lifeExp))
```
We can use the top_n() fucntion for counting the most extreme observation
```{r}
T1 %>% group_by(continent) %>% top_n(1,pop)
```


## Selecting and Transforming Data
We can use select() to select a range of columns or subtract a column.

We can use the contains arguement within select to get columns that contain a certain string. Other select helpers include starts_with(), ends_with(), last_col(), contains()
```{r}
 gapminder %>% select(country, -lifeExp, starts_with("Cont"),ends_with("op"))

```
We can use the rename() to rename a column name. We can also rename a column name inside the select() fucntion.
```{r}
 gapminder %>% select(country, -lifeExp, starts_with("Cont"),ends_with("op")) %>% rename(population=pop)

```
```{r}
 gapminder %>% select(country, -lifeExp, starts_with("Cont"),ends_with("op"), population =pop)

```

Transmute verb:

Transmute is a combination of select and mutate. It returns a subset of columns that are transformed and changed.
```{r}
T1 %>%
  transmute(country, continent, pop_inMillion = pop/1000000)
```


We should know that select() and rename() these verbs can't cahnge any variable while the mutate() and the transmute() can change the variables.


## Case Study: The babynames Dataset

We can filter for multiple observations or names using the filter(variable %in% c("",""))
```{r}
T1 %>% filter(continent %in% c("Ocenia","Asia"))
```
We can use ungroup() to ungroup the observations.
```{r}
T1 %>% filter(year==1952) %>% group_by(continent) %>% mutate(mean_lifeExp = mean(lifeExp)) %>% ungroup() %>% mutate(fraction = lifeExp / mean_lifeExp)
```
Window Function:

A window function takes a vector and returns another vector of the same length. We can use the lag() function.
```{r}
v <- c(10,30,55,123)
v
lag(v)
```

```{r}
gapminder %>% 
  select(country, year, pop) %>%
  filter(country == "India") %>%
  arrange(year) %>%
  mutate (difference_in_pop = pop - lag(pop)) %>%
  arrange((year))
```



