# Joining Data with dplyr

<https://learn.datacamp.com/courses/joining-data-with-dplyr>

```{r}
library(dplyr)
library(tidyverse)
library(gapminder)
library(lubridate)
```


## Joining Tables
Joining Tables is sometimes really important. We can use the inner_join() verb to join two datasets. But we need to make sure that the two tables we ar ejoining have one column in common and we have to join according to that column. 
```{r}
data1 <- iris%>% select(Species, Petal.Length,  Petal.Width) %>% mutate(index_1 =row_number())
data2 <- iris%>% select(Species, Sepal.Length,  Sepal.Width) %>% mutate(index_2 =row_number())
head(data1)
head(data2)
```

```{r}
join <- data1 %>% inner_join(data2,(by=c("index_1" = "index_2")))
join2 <- join %>% rename(index=index_1)
head(join2)
```
We can use suffix() to add any suffix to the same column names we are joining 
```{r}
data_joint <- data1 %>%
  inner_join(data2, by = c("index_1" = "index_2"), suffix = c("_main", "_notRequired")) %>% rename(index=index_1)

head(data_joint)
```
```{r}
main_data <- data_joint %>% select(Species=Species_main, Sepal.Length, Sepal.Width, Petal.Length, Petal.Width)
head(main_data)
```

We can also three or more tables using the inner_join() verb and pipe operators for more than once
```{r}
inventories <- readRDS("inventories.rds")
inventory_parts <- readRDS("inventory_parts.rds")
sets <- readRDS("sets.rds")
sets %>%
	# Add inventories using an inner join 
	inner_join(inventories, by = "set_num") %>%
	# Add inventory_parts using an inner join 
	inner_join(inventory_parts, by = c("id" = "inventory_id"))

```
We can also join any other table like this.

## Left and Right Joins
An inner join() only keep observations that appear in both tables.But if want to keep all the observations of the first table (left table) while joining we use the left_join verb. Similarly, if want to keep all the observations of the second table (right table) while joining we use the right_join verb.

```{r}
inventory_parts_joined <- inventories %>%
  inner_join(inventory_parts, by = c("id" = "inventory_id")) 
inventory_parts_joined
```

```{r}
# Aggregate Millennium Falcon for the total quantity in each part
millennium_falcon_colors <- inventory_parts_joined %>%
  filter(set_num == "7965-1") %>%
  group_by(color_id) %>%
  summarize(total_quantity = sum(quantity))

# Aggregate Star Destroyer for the total quantity in each part
star_destroyer_colors <- inventory_parts_joined %>%
  filter(set_num == "75190-1") %>%
  group_by(color_id) %>%
  summarize(total_quantity = sum(quantity))

# Left join the Millennium Falcon colors to the Star Destroyer colors
millennium_falcon_colors %>%
  left_join(star_destroyer_colors, by = "color_id", suffix = c("_falcon", "_star_destroyer"))
```
We can use the right_join() to keep all the observations in the second table including the matching observations in hte first table

```{r}
batmobile <- inventory_parts_joined %>%
  filter(set_num == "7784-1") %>%
  select(-set_num)

batwing <- inventory_parts_joined %>%
  filter(set_num == "70916-1") %>%
  select(-set_num)

batmobile
batwing
```


```{r}
batmobile %>%
  left_join(batwing, by = c("part_num", "color_id"), suffix = c("_batmobile", "_batwing"))
#Filter where Quantity_batmobile = NA
batmobile %>%
  right_join(batwing, by = c("part_num", "color_id"), suffix = c("_batmobile", "_batwing")) %>% 
  filter(is.na(quantity_batmobile))
#Replace NA with 0   
batmobile %>%
  right_join(batwing, by = c("part_num", "color_id"), suffix = c("_batmobile", "_batwing")) %>% 
  replace_na(list(quantity_batmobile = 0))
```

We can also join themes to themselves. 
```{r}
themes <- readRDS("themes.rds")
themes %>% 
    # Inner join the themes table
    inner_join(themes, by = c("id" = "parent_id"), suffix = c("_parent", "_child")) %>% 
    # Filter for the "Harry Potter" parent name 
    filter(name_parent == "Harry Potter")
themes %>% 
  # Left join the themes table to its own children
  left_join(themes, by = c("id" = "parent_id"), suffix = c("_parent", "_child")) %>%
  # Filter for themes that have no child themes
  filter(is.na(name_child))
```



## Full, Semi, and Anti Joins
We can join all the observations from both of the datasets using the full_join() verb
```{r}
batmobile %>%
full_join(batwing, by = c("part_num", "color_id"), suffix = c("_batmobile", "_batwing"))
```

There are two two type of filtering joins and the filtering joins doesn't add new variables rather it keeps or removes observation from the first table.

The semi_join() keeps the observations that are also in the first table it asks what observations are in X and also in Y 

```{r}
batmobile %>% 
  semi_join(batwing, by = c("color_id", "part_num"))
```
The anti join asks the question “What observations in X are NOT in Y?”

```{r}
batmobile %>% 
  anti_join(batwing, by = c("color_id", "part_num"))
```
Visualizing Set Differences :--  
We aggregate sets into colors 
```{r}
batmobile_colors <- batmobile %>% 
  group_by(color_id) %>%
  summarize(total = sum(quantity))
batwing_colors <- batwing %>% 
  group_by(color_id) %>%
  summarize(total = sum(quantity))
```

```{r}
colors <- readRDS("colors.rds")
#Comparinf Color Schemes of sets
colors_joint <- batmobile_colors %>% 
  full_join(batwing_colors, by = "color_id", suffix = c("_batmobile", "_batwing")) %>%
  replace_na(list(total_batmobile = 0, total_batwing = 0)) %>%
  #Adding the color names
  inner_join(colors, by = c("color_id" = "id")) %>% 
  #Adding fractions
  mutate(total_batmobile = total_batmobile / sum(total_batmobile),
         total_batwing = total_batwing / sum(total_batwing),
         difference = total_batmobile - total_batwing)
```
Now we can visualize the data 
```{r}
library(ggplot2)
library(forcats)

color_palette <- setNames(colors_joint$rgb, colors_joint$name)

colors_joint %>% 
  mutate(name = fct_reorder(name, difference)) %>% 
  ggplot(aes(name, difference, fill = name)) + 
  geom_col() + 
  coord_flip() + scale_fill_manual(values = color_palette, guide = FALSE)
```


## Case Study: Joins on Stack Overflow Data

The joining verbs :--  
*  inner_join()  
*  right_join()  
*  left_join()  
*  full_join()  
*  semi_join()   
*  anti_join()  

```{r}
question_tags <- readRDS("question_tags.rds")
questions <- readRDS("questions.rds")
tags <- readRDS("tags.rds")
answers <- readRDS("answers.rds")
```
```{r}
question_with_tags <- questions %>%
  inner_join(question_tags, by = c("id" = "question_id")) %>%
  inner_join(tags, by = c("tag_id" = "id"))

question_with_tags
#Most Common Tags 
commontags <- question_with_tags %>% count(tag_name, sort= TRUE)
commontags
```

```{r}
# Join the questions and question_tags tables
questions %>%
	left_join(question_tags, by = c("id" = "question_id")) %>%
# Join in the tags table
	left_join(tags, by = c("tag_id" = "id"))%>%
  # Replace the NAs in the tag_name column
    replace_na(list(tag_name = "only-r"))
```

```{r}
question_with_tags %>%
	# Group by tag_name
	group_by(tag_name) %>%
	# Get mean score and num_questions
	summarize(score = mean(score),
          	  num_questions = n()) %>%
	# Sort num_questions in descending order
	arrange(desc(num_questions))
```
Now we can add the answers with questions 
```{r}
questions %>%
	# Inner join questions and answers with proper suffixes
	inner_join(answers, by = c("id" = "question_id"), suffix = c("_question", "_answer")) %>%
	# Subtract creation_date_question from creation_date_answer to create gap
	mutate(gap = as.integer(creation_date_answer - creation_date_question)) 
```
The bind_rows() helps us to put one table on top of the other it's like stacking. 
```{r}


questions_type <- questions %>% mutate(type = "question")

answers_type <-  answers %>% mutate(type = "answer")

posts <- bind_rows(questions_type, answers_type)

#Creating date variable
questions_answers_year <- posts %>%
  mutate(year = year(creation_date)) %>%
  count(year, type)
#plotting
ggplot(questions_answers_year, aes(year,n, color = type)) +
  geom_line()
```
```{r}
# Inner join the question_tags and tags tables with the questions table
questions_with_tags <-questions %>%
  inner_join(question_tags, by = c("id" = "question_id")) %>%
  inner_join(tags, by = c("tag_id" = "id"))

# Inner join the question_tags and tags tables with the answers table
answers_with_tags <-answers %>%
  inner_join(question_tags, by = "question_id") %>%
  inner_join(tags, by = c("tag_id" = "id"))
# Combine the two tables into posts_with_tags
posts_with_tags <- bind_rows(questions_with_tags %>% mutate(type = "question"),
                              answers_with_tags %>% mutate(type = "answer"))

# Add a year column, then count by type, year, and tag_name
by_type_year_tag <-posts_with_tags %>%
  mutate(year = year(creation_date)) %>%
  count(type, year, tag_name)
# Filter for the dplyr and ggplot2 tag names 
by_type_year_tag_filtered <- by_type_year_tag %>%
  filter(tag_name %in% c("dplyr", "ggplot2"))

# Create a line plot faceted by the tag name 
ggplot(by_type_year_tag_filtered, aes(year, n, color = type)) +
  geom_line() +
  facet_wrap(~ tag_name)
```

